---
title: "Rational Analysis of Political Behavior"
subtitle: "Basic tools - Handout" 
author: "Jorge Fábrega"
date: "`r Sys.Date()`"
header-includes:
   - \usepackage{array}
output: 
  pdf_document:
    extra_dependencies: ["multirow"]
---

# Rationality, Strategy, and Equilibrium  

Strategy is the essence of politics. Examples:

*Executive Power*:  
o What agenda to follow?  
o Whom to appoint and in which position?  

*Legislative Power*:  
o Which candidates to present for the Supreme Court?  
o How to vote when the information is incomplete or there is no precedent?  

*Legislative Power*:  
o Which projects to support and which not?  
o What voting rules to establish?  
o Which legislative committee to join?  

*Political parties*:  
o Which candidates to run and where?  
o What alliances to form, when to break agreements?  

*Countries*:  
o With whom to negotiate agreements, treaties, etc., and on what terms?  
o When to threaten with or initiate a war?  

*Voters*:  
o For whom or what to vote in this election?  
o Why is strategy the essence of politics?  

*Two conditions*:  

o The "outcome of politics" does not depend solely on one's own action.  
o Each agent is interested in influencing that outcome according to their own purposes (rationality).  

# Unit of analysis: Interaction between individuals.  

Assumption: The structure (context) affects how each individual participates in that interaction.  

*Game Theory: A formal way to represent that interaction* that simplifies reality into a reduced set of elements:  
a) Identifies relevant players (strategic actors)  
b) For each player defines a set of actions they can take  
c) Defines a general decision-making criterion for each agent  
d) Defines a notion of strategic equilibrium.  

*Game Theory:*

a. DOES NOT SAY: "The world is like this."  
b. SAYS: "If it were like this, then agents should behave in this way," and then compares reality with the model.  

*Central interest topics in Political Science that can be treated as strategic games*:  

a. Role and implications of legislative rules.  
b. The deterrent power of weapons and diplomacy in international conflicts.  
c. Voting rules and participation in mass elections.  
d. Negotiation between strategic actors.  

*Why model?*  

a. There are no laboratories where all the variables affecting a political phenomenon are controlled.  
b. To elucidate what is necessary and sufficient:  
- i. Need to clarify assumptions about what is more important and what is peripheral.  
- ii. Sufficiency to explain a phenomenon of interest.  

*Rationality:*  

a. Assumption: Individuals have goals and the ability to choose between options to try to meet those goals.  

b. Rationality means that each individual will choose the best alternative to achieve that end.  
- i. The consistency of the choice matters.  
- ii. NOTE, we do not observe the goals, we deduce them from the actions of individuals.  

c. How do we formalize this simple idea?  
- i. We need to associate with each action an EXPECTED CONSEQUENCE, C. For example: Policies that the winning candidate implements after winning an election.  
- ii. To each actor PREFERENCES about those consequences:  
-- 1. C1 $\succeq$ C2: "C1 is at least as preferable as C2."  
-- 2. C1 $\sim$ C2: "C1 is at least as preferable as C2 and vice versa."  
-- 3. C1 $\succ$ C2: "C1 is more preferable than C2."  
- iii. We impose two structures on preferences:  
-- 1. COMPLETENESS: Every relevant pair of consequences (that is, every pair in C) can be compared   according to criteria R, I, or P.  
-- 2. TRANSITIVITY: For every trio of consequences C1, C2, and C3: If C1$\succeq$C2 and C2$\succeq$C3, then C1$\succeq$C3.  
- iv. GIVEN PREFERENCES: The set of preferences must be left fixed. This does not mean we believe people do not change tastes over time. This simply restricts the situations we can model. When the model fails to predict, we may be tempted to suggest that preferences changed, but we have to avoid falling into that temptation... Example:  
-- 1. If the model predicts that people should vote for X and they voted for Y, the modeler could say "people changed their minds"... but that is the same as saying, "I do not have a working predictive model."  
-- 2. This does not mean that the model must always predict the same decision. It means that it should predict changes in actions not changes in preferences.  
- v. ORDER OF PREFERENCES: With the previous assumptions, we can order preferences and assign them numbers... but only the order matters, not the magnitude of those numbers. That is, up to now:
{C1=5, C2=4, C3=3} is equal to {C1=15, C2=6, C3=-1000}.  

d. Here, rationality:  
- i. Is not a theory about how we learn to decide or form preferences. That process is not the subject of study here.  
- ii. Says nothing about the honorability of a preference. Only consistency matters. Classic example: Adolf Hitler.  
- iii. Does not imply that necessarily all rational actors will come to the same conclusion because:
-- 1. They may differ in goals.  
-- 2. They may face different risks and uncertainties or value those risks and uncertainties differently.  
-- 3. They may possess different information.  
- iv. Does not assume that rational actors are infallible.  

*Equilibrio*
La noción de equilibrio resume el objetivo de un análisis formal. Esto es: explicar la interacción.  Para nuestros fines equilibrio es (casi) sinónimo de PREDICCIÓN.

Partimos con “equilibrio” en decisiones individuales y luego veremos equilibrios propiamente estratégicos.

*Exercise: The Power of Deterrence*  
Examples:  
a. In international conflicts.  
- i. Economic sanctions  
- ii. Preemptive strike  

b. In the legislature  
- i. Oversight and calls to report to ministers  
- ii. Choosing a strong candidate so that the rival party refrains from presenting a candidate  
- iii. To introduce or not to introduce a bill or an amendment to a bill  

c. In citizen security  
- i. Exemplary punishment  
In a simple version of each of these situations, one player threatens another. The threatened party must decide whether to retaliate or not.  
a) What is the simplest set of actions available to the threatened?  
• Take the action, A  
• Withdraw, Not A  
b) What is the set of possible states?  
• Threatener carries out the threat, p  
• Threatener does not carry out the threat, 1-p  
c) What are the consequences/results?  
• Withdraws, U(R)  
• Does not withdraw and is punished, U(C)  
• Does not withdraw and is not punished, U(NC)  
d) Suppose that: U(NC) $\succeq$ U(R) $\succeq$ U(C)   
Let X1 and X2 be the two players in the deterrence game.  
X1 threatens X2 to punish him if he takes action A.  
X2 must decide between: A and Not A.  
There is a probability p that X1 will carry out his threat.  
What should X2 do?  

Utility of withdrawing: U(R)  
Utility of not withdrawing: $p*U(C) + (1-p)*U(NC)$  

X2 will be deterred if and only if: $U(R) > p*U(C) + (1-p)*U(NC)$  

Solving for p:  


$p > \frac{U(NC) - U(R)}{U(NC) - U(C)}$

How do we interpret the value of p?  
Answer: As the credibility of X1's threat.  

How do we interpret the right side of the inequality?  
Answer: As the level of critical risk faced. The numerator is the difference between going forward successfully and withdrawing, and the denominator is the difference between going forward successfully and unsuccessfully.  

\break 

# The technique and the art: Setting the problem for Chilean Politics in the 2010s  

All the above can be applied on real cases, but it requires some decisions to simplify the analysis. As an example consider some of the complexities that must to be solved for each aspect of the game.    

## 1. Players in the Game  

In the realm of Chilean politics during the 2010s, the key players can be identified through the lens of significant political events, such as electoral races, legislative reforms, or policy implementations. For instance, consider the case of education reform, a hot topic in Chile during this decade. 

Some players could be:  

$Government Officials$: Including the President, Ministers, and other elected officials who propose and support new policies.    
$Opposition Parties$: Political parties and leaders opposing the government’s stance, offering alternative solutions or critiques.  
$Student Unions and Educational Institutions$: Key stakeholders directly impacted by education reforms.  
$Public Opinion$: The collective sentiment of the populace which can influence policy decisions through protests, polls, or electoral support.  

In analyzing a specific scenario, one must delineate which players are actively influencing the outcome or problem under analysis and define their respective roles. Not all potential players contribute equally or relevant; hence, understanding the context helps in identifying the ones with actual decision-making power or influence.  

## 2. Actions Available to Players  

In the context of the same education reform, some relevant actions could be:  

Government Officials might propose new legislation, negotiate with other political entities, or engage in public relations campaigns to garner support.  

Opposition Parties could counter-propose alternative policies, mobilize public opinion against the government, or seek alliances to strengthen their position.  

Student Unions might organize protests, participate in public forums, or negotiate directly with policymakers.
  
Public Opinion influences through voting, participating in public discourse, or social media campaigns.  

Actions depend on the player’s position, resources, and objectives. In strategic analysis, it's crucial to map out possible actions for each player and understand how these actions interact within the political landscape.  

## 3. Players’ Preferences Among Alternatives  
Players’ preferences are shaped by their objectives, values, and the potential outcomes of their actions. For example:

Government Officials may prefer policies that ensure long-term educational improvement and political stability, balancing between reform and public approval.  
Opposition Parties prefer outcomes that enhance their political capital, potentially favoring alternatives that challenge the government's proposals while appealing to their electoral base.  
Student Unions prioritize reforms that directly address their demands for affordability, quality, and access to education.  
Public Opinion varies widely but generally leans towards equitable, quality education without undue financial burden.  
In developing a scenario, one must consider how different outcomes align with the preferences of each player, affecting their choices and strategies.  

## 4. Rules of the Game  
The rules in Chilean politics encompass at least the legal framework, institutional procedures, and unwritten norms governing player interactions. For education reform, this includes the legislative process, the role of the Ministry of Education, the budgetary constraints, and the mechanisms for public consultation and protest. Understanding these rules is vital for predicting how the game unfolds and how players can leverage different strategies within the constraints and opportunities the rules provide.  

## 5. Information Available to Players  
Information symmetry or asymmetry significantly influences decision-making. In our example:

Government Officials often have access to comprehensive data, expert analyses, and international benchmarks, guiding their policy proposals.  
Opposition Parties and Student Unions may have less access to official data but can leverage grassroots information, public sentiment, and academic research.  
Public Opinion is generally less informed but significantly influenced by media, social networks, and public figures.  
The dynamics of information flow, access, and utilization are critical in shaping the strategies and counter-strategies of the players involved.  

## Conclusion  
Analyzing a political scenario through the lens of game theory involves understanding the players, their possible actions, preferences, the governing rules, and the information landscape. By dissecting these components, one can better grasp the complexities of political decision-making and strategy. In the case of Chilean politics in the 2010s, the application of these game theory elements helps illuminate the multifaceted nature of policy debates, electoral strategies, and public engagements, providing a structured approach to analyzing political phenomena.  

\break

# The Median Voter Theorem  

The Median Voter Theorem (MVT) is a fundamental concept in the theory of electoral competition and public choice. It provides insight into how political outcomes are determined in democratic systems, especially under majority rule. The theorem assumes that individuals have preferences over outcomes, and these preferences can be ordered along a single-dimensional space, typically left to right on a political spectrum.  

## Key Elements of the Median Voter Theorem  

1. Players in the Game  
The primary players are the candidates competing for office and the voters who decide between them. In a two-candidate race, each candidate aims to secure a majority of votes.

2. Actions Available to Players  
Candidates can choose their position on a range of issues, represented as points along the political spectrum.  
Voters cast their votes based on which candidate’s position is closer to their own preferences.  

3. Players' Preferences Among Alternatives  
Voters have ideal points along the political spectrum where they are most satisfied. They prefer candidates whose policies are closest to these points.
Candidates seek to adopt positions that will win them the most votes. Under the MVT, this often means moving towards the median voter's position.

4. Rules of the Game  
The election is decided by majority rule, where the candidate with more than half of the votes wins.
Voters can only vote once, and they must choose one of the candidates.  

5. Information Available to Players
Candidates are aware of voters' preferences and distribution along the ideological spectrum.
Voters are informed about candidates' positions and can identify which candidate is closer to their own ideal point.

## Equilibrium and Assumptions  

Equilibrium: Under the Median Voter Theorem, the equilibrium occurs when both candidates position themselves at the median voter’s ideal point. This outcome is the Nash Equilibrium, where neither candidate can unilaterally change their position to gain more votes.  

Assumptions:  
*Single-peaked Preferences*: Voters’ preferences are single-peaked, meaning they have one ideal point, and their satisfaction decreases as policies move away from this point.  
*Unidimensional Issue Space*: All political issues can be represented on a single left-right spectrum.  
*Full Information*: Voters and candidates have full knowledge of each other's preferences and positions.  
*Majority Rule*: The candidate with the majority of votes wins.  
*No Indifference*: Voters will always prefer the candidate whose position is closer to their ideal point.  

## Application and Implications  

The Median Voter Theorem has significant implications for political strategy, as it predicts that in a two-party system, parties or candidates will gravitate toward the center to capture the median voter.
It also suggests that policies enacted will reflect the preferences of the median voter, which can lead to centrist policies in politically balanced populations.  

## Conclusion  

The Median Voter Theorem offers a powerful tool for understanding electoral outcomes and policy decisions. While actual political behavior may deviate due to multi-dimensional issues, strategic voting, or imperfect information, the MVT provides a baseline from which to analyze political strategies and outcomes. In the context of Chilean politics in the 2010s, applying these game theory elements helps illuminate the strategic considerations of candidates and the centrality of moderate voters in determining election results and policy directions.  

\break  

# Exercise 1: Scheduling Study Group Sessions

## Scenario:
A university study group of seven members needs to decide on which day of the week they will hold their regular study sessions for the upcoming semester. The decision is made by majority vote, and if no majority is reached, the study session will default to the day it was held the previous semester, which was Thursday.  

Each member has a preferred day, and their preferences decrease linearly from this peak as the days move away from their ideal choice. The members have the following ideal days:  

Member 1: Monday  
Member 2: Tuesday  
Member 3: Wednesday  
Member 4: Thursday  
Member 5: Friday  
Member 6: Saturday  
Member 7: Sunday  

## Questions:
a) If Member 3 initiates the voting process, what day is most likely to be chosen for the study sessions according to the Median Voter Theorem? Explain your reasoning, considering that individuals prefer days closer to their ideal.  
b) Now imagine the university introduces a new policy allowing groups to meet only on weekdays. With this constraint, how does the result change? Use the Median Voter Theorem to determine the new most likely meeting day.  

# Exercise 2: Allocating Budget for Neighborhood Pathway

## Scenario:  
The neighborhood council is allocating a portion of its annual budget to improve pathways in the community. The council needs to decide what percentage of the total budget should be dedicated to this project. The decision is made through a majority vote among the seven council members. They must choose a percentage between 0% and 100%, and they prefer allocations closer to their ideal percentages. If no consensus is reached, the budget allocation will remain the same as last year, which was 30%.  

Each council member has a preferred budget allocation for the pathway, as follows:  

Member A: 10%  
Member B: 20%  
Member C: 40%  
Member D: 50%  
Member E: 60%  
Member F: 70%  
Member G: 80%  

## Questions:  
a) If Member D leads the discussion and sets the initial proposal, what budget allocation percentage is likely to be chosen according to the Median Voter Theorem? Explain your reasoning based on the assumption that each member's satisfaction decreases as the chosen percentage moves away from their ideal.  
b) Assuming that the council agrees to focus solely on infrastructure improvements this year and must allocate a minimum of 25% of the budget to the pathway due to contractual obligations, how would this constraint affect the final decision? What would the Median Voter Theorem predict as the new budget allocation percentage?  

\break

# The McKelvey Chaos Theorem

The McKelvey Chaos Theorem, developed by political scientist Richard D. McKelvey in the field of social choice theory, addresses the instability inherent in multidimensional vote spaces. While the Median Voter Theorem provides a clear predictive outcome in one-dimensional political spaces, McKelvey's theorem highlights the unpredictability that can arise when decisions are based on multiple dimensions or issues.  

## Key Points of the McKelvey Chaos Theorem:

*Multidimensional Policy Spaces*: The theorem applies to situations where policy choices are made based on more than one issue, leading to a multidimensional space of possible outcomes.  

*Cycling Majority Preferences*: In multidimensional spaces, it is possible for preferences to cycle indefinitely. This means that for any possible policy, it is possible to find another policy that a majority prefers. This cycle can continue without reaching a stable outcome, leading to "chaos" in collective decision-making.  

*Absence of a Core*: Unlike in one-dimensional spaces, where the median voter's preference can act as a stable equilibrium (the "core"), multidimensional spaces may lack such a core. Without a core, there is no stable policy outcome that cannot be defeated by another through majority voting.  

*Manipulation by Agenda Setters*: Given the lack of stability, individuals or groups who control the order in which decisions are made (agenda setters) can significantly influence outcomes. They can structure the sequence of votes to lead to an outcome they prefer, even if it is not a majority's first choice.  

## Example of the McKelvey Chaos Theorem:  

Consider a city council deciding on the allocation of a budget across three different areas: Education (E), Infrastructure (I), and Public Safety (S). Each council member (A, B, C, D, E) has different priorities:

Member A: E > I > S  
Member B: I > S > E  
Member C: S > E > I  
Member D: E > S > I  
Member E: S > I > E  

In this multidimensional setting (where preferences across three areas can be considered), no single allocation becomes a stable choice. For instance:  

If the council initially supports more budget for Education, a majority might prefer shifting some funds to Infrastructure (A and B prefer I over E). From Infrastructure, a new majority (C and E) might shift the budget preference to Public Safety. But with a focus on Public Safety, a different majority (A and D) could push back towards Education. This cycle could continue indefinitely with no stable outcome, exemplifying the "chaos" as described by McKelvey. Each member's preference leads to a cycle without a clear majority for any single policy.  

Given this chaos, if Member A were the agenda setter, they could strategically arrange votes to move from Public Safety to Education, ensuring the final allocation is closest to their top preference, showcasing how an agenda setter can manipulate outcomes in a multidimensional setting.  

# Exercise 1: Chilean Constitutional Reform  

## Scenario:  
In the 2020s, Chile faces a pivotal moment in deciding on constitutional reforms. The reform process involves multiple dimensions, including governance structure, social rights, environmental regulations, and indigenous representation. A constitutional assembly, comprised of members from various political and social backgrounds, must reach a decision on these issues. Each member prioritizes different aspects of the reform based on their political affiliation and constituency demands.  

Preferences of the assembly members (simplified for this exercise) are as follows:  

Member 1 (Left-wing): Social rights > Environmental regulations > Indigenous representation > Governance structure  
Member 2 (Right-wing): Governance structure > Economic regulations > Social rights > Indigenous representation  
Member 3 (Environmental group): Environmental regulations > Social rights > Indigenous representation > Governance structure  
Member 4 (Indigenous representative): Indigenous representation > Environmental regulations > Social rights > Governance structure  
Member 5 (Centrist): Social rights > Governance structure > Environmental regulations > Indigenous representation  

## Questions:  
a) Given these multidimensional preferences, demonstrate how the McKelvey Chaos Theorem applies to the decision-making process in the assembly. What challenges arise in reaching a stable consensus on the constitutional reform?  
b) Assume Member 5 becomes the agenda-setter. How might they manipulate the order of decisions to achieve an outcome closer to their preferences? Construct a sequence of votes that demonstrates this manipulation, considering the McKelvey Chaos Theorem.  

# Exercise 2: Chilean Pension System Reform  

## Scenario:
Chile's pension system has been under scrutiny, and various political factions propose different reforms. The debates encompass several dimensions: pension fund contributions, benefit levels, state involvement, and investment regulations. The national congress, consisting of representatives from different political parties, must come to an agreement on these dimensions.  

The representatives have the following order of preferences:  

Representative A (Progressive party): State involvement > Benefit levels > Pension fund contributions > Investment regulations  
Representative B (Liberal party): Investment regulations > Pension fund contributions > Benefit levels > State involvement  
Representative C (Conservative party): Pension fund contributions > Investment regulations > State involvement > Benefit levels  
Representative D (Socialist party): Benefit levels > State involvement > Investment regulations > Pension fund contributions  
Representative E (Green party): State involvement > Environmental sustainability in investments > Benefit levels > Pension fund contributions  

## Questions:  
a) Illustrate how the McKelvey Chaos Theorem could manifest in the congress’s attempts to reform the pension system. What cyclical patterns could emerge based on the representatives’ preferences?  
b) If Representative C is given the power to set the agenda, how could they strategically arrange the voting process to end up with a reform package that aligns closely with the Conservative party's preferences? Describe the voting order and expected outcomes based on the principles of the McKelvey Chaos Theorem.  

\break

# Some additional exercises to find Nash equilibria  

Find the Nash Equilibrium or Equilibria in each of the following games.

The Nash Equilibrium is a fundamental concept in game theory, particularly in the analysis of competitive environments. It describes a situation where, given the strategy choices of the other player, no player has anything to gain by changing their own strategy unilaterally. This equilibrium provides a stable set of expectations for players, as each player's strategy is the best response to the strategies of others.

## A Simple Case:  
In this example, we delve into a strategic interaction involving two players, each faced with four possible choices. The players, denoted as Player 1 and Player 2, must decide independently and simultaneously without knowing the choice of the other. The outcome of each combination of choices is represented as a payoff matrix, where the entries show the payoffs to each player under the various scenarios.  

In the payoff matrix provided below, each cell contains two numbers: the first number represents the payoff to Player 1, and the second number represents the payoff to Player 2, given the strategies chosen by both. The choices made by Player 1 are represented by the rows of the matrix, while the choices made by Player 2 are represented by the columns.  

The goal here is to identify the Nash Equilibrium or Equilibria within this setting. To find these equilibria, we need to consider what each player will choose, knowing the potential choices of their opponent and aiming to maximize their own payoff.  

Let’s inspect the strategic landscape as laid out in the following payoff matrix and determine the conditions under which a Nash Equilibrium is achieved. Understanding the rationale behind the players' choices in this matrix can provide insights into strategic behavior in real-world scenarios, such as competitive markets, political campaigns, or negotiations.  

Now, observe the matrix below and identify the Nash Equilibria. Reflect on how the decision of one player influences the optimal choice of the other and vice versa, leading to points in the matrix where both players are making their best possible moves in response to one another.  
```{r table1, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
# Required Libraries
library(knitr)
library(kableExtra)

# Create the data frame
# Corrected Matrix Creation in R
tableData <- matrix(c('1, 10', '3, 14', '3, 1', '2, 6',
                      '2, 15', '4, 8', '2, 4', '5, 9',
                      '-1, 3', '0, 4', '3, 5', '9, 0',
                      '4, 9', '4, 8', '1, 1', '7, 10'), 
                    nrow = 4, byrow = TRUE, 
                    dimnames = list(c("A", "B", "C", "D"), c("A", "B", "C", "D")))



# # Print the table with kable and kableExtra for formatting
kable(tableData, "latex", booktabs = TRUE, caption = "Strategy Payoffs", align = 'c') %>%
  kable_styling(latex_options = c("striped", "scale_down"), font_size = 10) %>%
  add_header_above(c("Player 1" = 1, "Player 2" = 4), escape = FALSE) %>%
  column_spec(1, border_right = TRUE)

```

\newpage    


## Urban Development Projects

In the realm of urban development, city councils are pivotal in shaping the landscape and well-being of the community. They face decisions on projects that impact urban growth, environmental health, and residents' quality of life. In this scenario, we focus on a city council (Player 1) evaluating four development projects, and the community (Player 2), which is affected by these decisions.  

Development Projects under Consideration:  

Park (P): Aims to create recreational green spaces, enhancing environmental quality and community life.  
Residential Area (R): Seeks to address housing shortages, supporting population growth and local economy.  
Commercial Zone (C): Designed to stimulate economic development by attracting businesses and creating jobs.  
Industrial Park (I): Focuses on industrial growth, aiming to provide employment opportunities and bolster the city’s economic base.  

**Stakeholders**:  

Local Government (Player 1): Prioritizes projects based on economic returns, growth potential, and sustainability.  
Community (Player 2): Values quality of life, environmental health, and access to services.  

The upcoming decision involves complex trade-offs among economic, environmental, and social factors. The payoff matrix illustrates the benefits and costs (in millions) for both the local government and the community, reflecting the outcomes of each project.  

This scenario underscores the application of game theory in urban planning, where the city council must align its strategy with the community's needs and preferences. The goal is to find a balance that maximizes mutual benefits and addresses the pressing concerns of urban development.  

Let's delve into the payoff matrix to evaluate which urban development project aligns best with the interests of both the local government and the community, aiming to uncover a strategic path that fosters sustainable urban growth and community welfare.   

```{r table2, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
# R code to create the matrix for Urban Development Projects
urbanDevProjects <- matrix(c('5, 12', '2, 8', '4, 10', '1, 5',
                             '6, 9', '7, 15', '3, 7', 'NA, NA',
                             '4, 11', '3, 13', '5, 14', 'NA, NA'), 
                           nrow = 3, byrow = TRUE, 
                           dimnames = list(c("Park", "Residential", "Commercial"),
                                           c("LocalGov", "Community", "Business", "Environment")))

# # Print the table with kable and kableExtra for formatting
kable(urbanDevProjects, "latex", booktabs = TRUE, caption = "Strategy Payoffs", align = 'c') %>%
  kable_styling(latex_options = c("striped", "scale_down"), font_size = 10) %>%
  add_header_above(c(" " = 1, "Player 2" = 4), escape = FALSE) %>%
  column_spec(1, border_right = TRUE)

```

\newpage  

## Public Health Initiatives  

The landscape of public health is complex, requiring strategic decisions that significantly impact the community's welfare. Health departments (Player 1) must navigate these decisions, implementing initiatives that address both current challenges and future community health needs (Player 2).  

Key Initiatives Under Consideration:  

Vaccination Campaign (VC): Targets increasing community vaccination rates to prevent diseases.  
Public Awareness Campaign (PA): Aims to educate the public on health issues for better individual decision-making.  
Improved Hospital Facilities (IH): Seeks to enhance healthcare services for improved patient outcomes.  
Emergency Response Teams (ER): Prepares specialized units for quick responses to public health emergencies.  

**Players**:

Health Department (Player 1): Responsible for evaluating and implementing health initiatives based on efficiency, impact, and resource allocation.  
Community (Player 2): The direct recipient of health initiatives, whose participation and response significantly influence the initiatives' success.  
The decision-making process involves assessing each initiative's benefits against its challenges and costs, considering the community's needs and the department's capacity. The provided payoff matrix visualizes these aspects, offering a comparative analysis of the initiatives' potential health outcomes and community support levels.  

This scenario underscores the ethical and social considerations inherent in public health decisions. The health department must balance immediate public health improvements with sustainable, long-term benefits to the community. The forthcoming analysis of the payoff matrix aims to identify which public health strategy will most effectively enhance the community's well-being, aligning with the department's goals and the community's interests.  

Next, we will explore the payoff matrix for these Public Health Initiatives, aiming to pinpoint the approach that ensures the best health outcomes for the community while aligning with the department's strategic objectives.   

```{r table3, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
# R code to create the matrix for Public Health Initiatives
publicHealthInitiatives <- matrix(c('8, 13', '5, 10', '7, 9', '9, 15',
                                    '6, 11', '7, 12', 'NA, NA', '8, 14',
                                    '4, 8', '3, 6', 'NA, NA', '5, 7'), 
                                  nrow = 3, byrow = TRUE, 
                                  dimnames = list(c("Vaccination", "PublicAwareness", "HospitalFacilities"),
                                                  c("Effectiveness", "PublicSupport", "Cost", "Speed")))


# # Print the table with kable and kableExtra for formatting
kable(publicHealthInitiatives, "latex", booktabs = TRUE, caption = "Strategy Payoffs", align = 'c') %>%
  kable_styling(latex_options = c("striped", "scale_down"), font_size = 10) %>%
  add_header_above(c(" " = 1, "Player 2" = 4), escape = FALSE) %>%
  column_spec(1, border_right = TRUE)
```

\newpage  
  
## Environmental Policy Decisions  
Public Health Initiatives: Strategic Choices
The realm of public health is dynamic, with health departments (Player 1) orchestrating initiatives to combat health challenges while promoting overall community health (Player 2). These departments must make strategic decisions that significantly affect public welfare and community resilience against health threats.

Public Health Initiatives:  

Vaccination Campaign (VC): Aims to boost community immunity through increased vaccination rates.  
Public Awareness Campaign (PA): Focuses on educating the public on health practices and disease prevention.  
Improved Hospital Facilities (IH): Involves enhancing healthcare infrastructure for better patient care.  
Emergency Response Teams (ER): Develops specialized units for swift responses to health crises.  

**Players**:

Health Department (Player 1): Evaluates and selects public health initiatives based on potential impacts, resource allocation, and ethical considerations.  
Community (Player 2): The beneficiary and participant of health initiatives, whose engagement and support are vital for the success of chosen strategies.  
In this scenario, the health department is faced with the challenge of selecting the most appropriate initiative. The department must balance health outcomes with practical constraints and community expectations. The provided payoff matrix elucidates the trade-offs between initiative effectiveness and public approval.  

The decision carries profound implications, shaping the health landscape and reflecting the department's commitment to public welfare. In evaluating the options, the health department considers both short-term benefits and long-term health advancements, aiming to align its choice with community needs and values.  

We will next explore the payoff matrix to identify the most suitable public health strategy, considering the department's goals (Player 1) and the community's well-being (Player 2).  


```{r table4, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
# R code to create the matrix for Environmental Policy Decisions
environmentalPolicy <- matrix(c('9, 20', '3, 12', '6, 15', '8, 25',
                                '7, 18', '5, 10', '4, 8', 'NA, NA',
                                '5, 14', '6, 16', '2, 5', '7, 22',
                                'NA, NA', 'NA, NA', 'NA, NA', '5, 19'), 
                              nrow = 4, byrow = TRUE, 
                              dimnames = list(c("EmissionReduction", "Reforestation", "PollutionControl", "GreenTech"),
                                              c("Impact", "EconomicCost", "PublicApproval", "Longevity")))


# # Print the table with kable and kableExtra for formatting
kable(environmentalPolicy, "latex", booktabs = TRUE, caption = "Strategy Payoffs", align = 'c') %>%
  kable_styling(latex_options = c("striped", "scale_down"), font_size = 10) %>%
  add_header_above(c(" " = 1, "Player 2" = 4), escape = FALSE) %>%
  column_spec(1, border_right = TRUE)
```

